
---
title: "Using R to Search for future mining and landing sites on Mars"
author: Hadarou Sare

---

# Introduction

For more than decades,NASA and its partners has been developing technologies to extract precious metals and water from the Moon, Mars and other bodies in the solar system. The extraction of water is very important because it can reduce considerably the cost of launch since the water can be split into Hydrogen and Oxygen and use as propellant in space. That idea has been clearly supported recently by the US goverment and European partners with the Artemis and other projects. However, even though scientists and engineers have already developed the right technologies able to extract water and other resources from space, it is still unclear where they should land their technologies(Rover and others) and start mining the resources they are looking for. 

The goal of this project is to use the package "threejs" to visualize the future best mining sites on Mars(water/water ice/volatiles materials mining sites). The final product is a map showing locations of best landing sites and mining sites.
However, before getting into that step, I used some geological and engineering criteria to determine and selected the best landing/mining sites for the future rovers. The criteria used to select the landing/mining sites are: where are located minerals indicating the presence of water/water ice?, where do we have permanent sunlight to use as source of energy for processing the water and charging the rovers? Is the topography of the site suitable for the mobility of the rover to and from the processing center?... 
For this project, the most important criteria I  considered is the geological criteria for question of time. I choosed and explored 9 sites as training sites and then selected among those sites the best ones as future landing/mining sites.

# Materials and methods

I choosed 9 sites as training sites.
For each training site, I downloaded and used Raster data of  Mars. Those data are Visible and infrared data in the “.TIFF” format and are combination of different band (7 bands according to Mars Odysee website) and each band is showing data in a specific wavelength. After downloading those data, I  resized some of them so that all images will have the same size and I  then added them together to create a metadata containing all information from each site. I finally used the band ratio combination to determine the presence and abundance of water/water ice, ice or volatiles material in each site I choose as training site and from there I was able to see which sites could be selected as the best future landing/mining sites.

# Data and Results
## PART1: Selection of thhe landing/mining sites

The goal here is to detect the site where we could have more water/ water ice/ volatiles materials. I applied some corrections to satelitte images I downloaded. For question of time I will only show what I did in only one site here but all sites I selected to be future landing/mining site are shown later in the second part of this project.

I loaded required packages

```{r, message=F, warning=F}
library(raster)
library(tidyverse)
library(sf)
library(rpart)
library(rpart.plot)
library(rasterVis)
library(mapedit)
library(mapview)
library(caret)
library(forcats)
library(sp)
library(raster)
```

I prepare the data by reading in the different bands that comprise the satelitte imagery. Each band refers to different spectrum.


```{r, message=F, warning=F}
band1 <- ("B02.tif")
band2 <- ("B03.tif")
band3 <- ("B04.tif")
band4 <- ("B05.tif")
band5 <- ("B06.tif")
band6 <- ("B07.tif")
band7 <- ("B08.tif")
```



```{r, message=F, warning=F}
library(raster)
band1 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/B02.tif")
#(band1 <- raster("B02.tif"))
band2 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/B03.tif")
band3 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/B04.tif")
band4 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/B05.tif")
band5 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/B06.tif")
band6 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/B07.tif")
band7 <- raster("C:/Users/Owner/Desktop/sare  backup/c/user hadar/Desktop/GEO511/Tasks/geo511-tasks-hsare/geo511-2020-final-project/geo511-2020-project-hsare/selection_of_landingMining_sites/B08.tif")
```

Show result of band1

```{r}
print(band1)
```

Now, I check the size of image before combining them together using STACK to make sure the sizes match. The code below shows that band1, band2, band3, and band7 are 10x10. Band4, band5, and band6 are 20x20.


```{r}
res(band1) # Size find is 10x10
res(band2) # Size find is 10x10
res(band3) # Size find is 10x10
res(band4) # Size find is 20x20
res(band5) # Size find is 20x20
res(band6) # Size find is 20x20
res(band7) # Size find is 10x10
```

I resize now some image to make sure all image have the same size before adding them together.
From the results above, I then multiply each image (band1, band2, band3, and band7) by 2 to have 20x20.


```{r}
band1 <- aggregate(band1, fact = 2) # multiply the size of band1 by 2 to have 20x20
band2 <- aggregate(band1, fact = 2) # multiply the size of band2 by 2 to have 20x20
band3 <- aggregate(band1, fact = 2) # multiply the size of band3 by 2 to have 20x20
band7 <- aggregate(band1, fact = 2) # multiply the size of band7 by 2 to have 20x20
```

Now I can add all bands together to have a metadata containing all informations I am looking for and which will be helpful to continue the project.

I decided to add only band2, band3 and band7 together because those wavelength are able to reflect the distribution of water on the sites. Also I just used those bands because the code does not work when I add the remaining band to the current I already stack together.
```{r}
image <- stack(band2, band3, band7)
#image1 <- stack(band4, band5, band6)

```

Let's explore the images. Here, we are able to see the number of band used, the coordinate system the imagery is projected in, the resolution (or grid cell size) of the raster, and the resolution of the images


```{r}
nlayers(image) #This give the number of bands use
crs(image) #the coordinate system the imagery is projected in, and the resolution (or grid cell size) of the raster.
res #resolution of the images
```

##Results1
Now that we know a little more about the imagery we are using, let plot it. Since image is a multi-band raster,
we use the plotRGB function from the raster package, which allows us to specify what bands should be visualized.

###TRUE COLOR PLOTS. It uses the red band (4) for red, the green band (3) for green, and the blue band (2) for blue.

t
```{r}
par(col.axis="white",col.lab="white",tck=0)
plotRGB(image, r = 4, g = 3, b = 2, axes = TRUE, 
        stretch = "lin", main = "True Color Composite")
box(col="white")
```
###FALSE COLOR PLOT.The false color composite uses NIR (5) for red, red (4) for green, and green (3) for blue.


```{r}
par(col.axis="white",col.lab="white",tck=0)
plotRGB(image, r = 5, g = 4, b = 3, axes = TRUE, stretch = "lin", main = "False Color Composite")
box(col="white")
```
###Indices of WATER

We can use Modified Normalized Difference Water Index [MNDWI=(Green-MIR)/ (Green+MIR)] to separate water from land. This idea is from Reference=https://www.researchgate.net/post/What_is_the_best_band_combination_for_highlighting_the_water_and_soil_together_in_LANDSAT_image2
The Author of the reference is Saygin Abdikan and is currently working at the Department of Geomatics Engineering, Hacettepe University. The Author researchGate is https://www.researchgate.net/profile/Saygin_Abdikan

The formula is as below:
[MNDWI=(Green-MIR)/ (Green+MIR)]
MNDWI <- (image[[2]] - image[[3]])/(image[[2]] + image[[3]]).

To plot the results with ggplot, we convert the raster into a data frame and use geom_tile.
The plot generated by the codes below show where are located water/ water ice/ volatiles materials the most and we have then know where to land the rovers and start the mining process.

Band4, band5, and band6 are use here. Band4 represent the Green band in our case and band6 represent the MIR band


```{r}
#MNDWI <- (image[[2]] - image[[3]])/(image[[2]] + image[[3]])
#MNDWI <- ((band2 + band3) - band7)/((band2 + band3) + band7)
MNDWI <- (band4 - band6)/(band4 + band6)


as(MNDWI, "SpatialPixelsDataFrame") %>% 
  as.data.frame() %>%
  ggplot(data = .) +
  geom_tile(aes(x = x, y = y, fill = layer)) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        panel.grid.minor = element_blank()) +
  labs(title = "Water or Water ice or Volatiles materials", 
       x = " ", 
       y = " ") +
  scale_fill_gradient(high = "#CEE50E", 
                      low = "#087F28",
                      name = "Water or Water ice or Volatiles materials")
```

## PART1: Visualization of landing/mining sites on map

I loaded required packages

```{r, message=F, warning=F}
library(threejs)
```

##Results2
I prepared and loaded the data.


#I decided not to load this data because my codes are not working yet
#```{r}


#the_martian <- read.csv("the_martian.csv", comment.char="#")


#globejs(img="Mars_Viking_MDIM21_ClrMosaic_global_1024.jpg", bodycolor="#555555", emissive="#444444", 
#        lightcolor="#555555", 
#        lat=the_martian$x, 
#        long=the_martian$y)
#```



# Conclusions

Using R , we are able to resize the images, correct the images, and process the images to detect the best future landing and mining sites. This project can be extended to other area and could help scientists and researcher in selecting not only landing and mining sites but also detecting and forcasting temperature in space, maping minerals in space, detecting geological features in space.
[~200 words]


# References

All sources are cited (to be updated later)
